---
title: "SarcomaPreOpRTx"
author: "Heejoo"
date: "`r Sys.Date()`"
output:
  html_document: 
    dev: jpeg
    df_print: paged
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: no
    code_folding: show
  word_document:
    toc: yes
    toc_depth: '3'
  bookdown::word_document2:
    fig_caption: yes
editor_options:
  chunk_output_type: console
always_allow_html: yes
---

<style>
   p {line-height: 2.5em;}
   pre, code {white-space:pre !important; overflow-x:scroll !important}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

## Load & Input
```{r}
library(readxl)
library(dplyr)

#read data

setwd("C:/Users/USER/Desktop/2020-1학기/의학통계/sarcoma/preOP RTx")

## 읽을 때부터 UK를 결측치로 
a <- excel_sheets("sarcoma data sheet SMC 20200716.xlsx") %>% 
  lapply(function(x){read_excel("sarcoma data sheet SMC 20200716.xlsx",sheet=x, skip=2, na = c("UK"))})
a.rps<-excel_sheets("preop Bx. RPS 여부 20200712.xlsx") %>% 
  lapply(function(x){read_excel("preop Bx. RPS 여부 20200712.xlsx",sheet=x, skip=2, na = c("UK"))})

b <- a[[1]] %>% 
  left_join(a[[2]], by = "환자번호") %>% left_join(a[[3]], by = "환자번호") %>% left_join(a[[4]], by = "환자번호") %>%
  left_join(a[[5]], by = "환자번호") %>% left_join(a[[6]], by = "환자번호") %>% left_join(a[[7]], by = "환자번호")

#Age 계산
b$Age <- as.numeric(b[["수술날짜\r\n\r\ndd-mm-yyyy"]] - b[["생년월일\r\n\r\ndd-mm-yyyy"]])/365.25
```

## Methods
```{r}
# c <- b %>% 
#   filter(`Primary 수술여부\r\n\r\n0. Primary tumor\r\n1. Residual after incomplete resection\r\n2. Local recurrence.x`== 0)
c<-b

out <- c %>% select(환자번호,Age,`성별\r\n\r\nM/F`)
names(out)[3] <- "Sex"; names(out)[1] <- "ID"
out$Sex <- as.factor(out$Sex)

out.rps<-a.rps[[1]][c(4:6)]
names(out.rps)[1]<-"ID"; names(out.rps)[2]<-"preOP Retroperitoneal sarcoma (RPS)"; names(out.rps)[3]<-"postOP Retroperitoneal sarcoma (RPS)"
out<-merge(out,out.rps,by="ID")


#preOpRTx : 시행(1|5)=TRUE, 시행x(4)=FALSE, 그외=NA
out$preOpRTx <- ifelse(c[["RT timing\r\n\r\n0.None \r\n1.Preop only\r\n2. IORT only\r\n3.Preop + IORT\r\n4.Postop only\r\n5.Preop + postop boost\r\n6.IORT + postop"]] %in% c("1", "5"), 1,
                       ifelse(c[["RT timing\r\n\r\n0.None \r\n1.Preop only\r\n2. IORT only\r\n3.Preop + IORT\r\n4.Postop only\r\n5.Preop + postop boost\r\n6.IORT + postop"]] == "4", 0, NA))

#TE : 삽입(1)=TRUE, 삽입x(0)=FALSE
out$TE <- ifelse(c[["Tisuue expander insertion \r\n유뮤\r\n\r\n0. No\r\n1. Yes"]] == "1", 1,
                 ifelse(c[["Tisuue expander insertion \r\n유뮤\r\n\r\n0. No\r\n1. Yes"]]== "0",0, NA))

#Group
#"1" : preOP RTx 시행o
#"2" : preOP PTx 시행x, TE 삽입 x
#"3" : preOP RTx 시행x, TE 삽입 o
out$Group <- as.factor(ifelse(out$preOpRTx==1,1,ifelse(out$TE ==0,2,3)))
out$Group1vs23 <- as.factor(ifelse(out$Group == 1,1,23))
out$Group1vs2 <- as.factor(ifelse(out$preOpRTx==1,1,ifelse(out$TE ==0,2,NA)))

out
```

Data for Table1
```{r}
#BMI
height <- as.numeric(c[["키\r\n(cm)"]])/100
weight <- as.numeric(c[["몸무게\r\n(kg)"]])
out$BMI <- weight/height/height
out$BMI_cat <- factor(ifelse(out$BMI < 18.5, "< 18.5", ifelse(out$BMI < 25, "< 25", ifelse(out$BMI < 30, "< 30", "≥ 30"))))


#DM : 1=TRUE, 0=FALSE
out$DM <- as.integer(c[["DM\r\n\r\n0. No\r\n1.yes"]])

#HTN : 1=TRUE, 0=FALSE
out$HTN <- as.integer(c[["HTN\r\n\r\n0. No\r\n1.yes"]])

#COPD : 1=TRUE, 0=FALSE
out$COPD <- as.integer(c[["COPD\r\n\r\n0. No\r\n1.yes"]])

#Coronary artery disease : 1=TRUE, 0=FALSE
out$CoronaryArteryDisease <- as.integer(c[["Coronary artery disease\r\n\r\n0. No\r\n1.yes"]])

#Chronic renal disease : 1=TRUE, 0=FALSE
out$ChronicRenalDisease <- as.integer(c[["Chronic renal disease\r\n\r\n0. No\r\n1.yes"]])


#Prev abdominal OP Hx : 0="무", 1="유", 2="기타(laparo)"
out$PrevAbdominalOp <- c[["이전\r\nabdominal op Hx \r\n여부\r\n\r\n0. 무\r\n1. 유\r\n2. 기타(laparo)"]]
#out$PrevAbdominalOp <- as.factor(out$PrevAbdominalOp)

#PreOP chemo : 1=TRUE, 0=FALSE
out$preOpChemo <- as.integer(c[["Neoadjuvant chemo 여부\r\n\r\n0.No\r\n1.Yes"]])

#Hb
out$Hb <- as.numeric(c[["수술전 \r\n피검사\r\n\r\nHb\r\n(g/dL)"]])
out$Hb_below9 <- as.integer(out$Hb < 9)  
out$Hb_below10 <- as.integer(out$Hb < 10)
  

#Albumin
out$Albumin <- as.numeric(c[["수술전 피검사\r\n\r\nAlbumin\r\n(g/dL)"]])
out$Albumin_below3 <- as.integer(out$Albumin < 3)

#PLT
out$PLT <- as.numeric(c[["수술전 피검사\r\n\r\nPlatelet\r\n(1000/uL)"]])
out$PLT_below50 <- as.integer(out$PLT < 50)
out$PLT_below100 <- as.integer(out$PLT < 100)

#PT INR
out$PT_INR <- as.numeric(c[["수술전 피검사\r\n\r\nPT(INR)"]])
out$PT_INR_over1.5 <- as.integer(out$PT_INR > 1.5)

#Primary or Recurrent tumor : 0=Primary, 1:Residual after incomplete resection, 2=local recurrence
out$PrimaryRecurrent <- as.integer(c[["Primary 수술여부\r\n\r\n0. Primary tumor\r\n1. Residual after incomplete resection\r\n2. Local recurrence.x"]])

#Tumor size
out$TumorSize <- as.numeric(c[["종양크기\r\nFirst dimension\r\n(mm)"]])

#Tumor histologic subtype
#LPS : 0. WD Liposarcoma / 1. DD Liposarcoma / 2. Pleomorphic Liposarcoma / 7. 중 comment 에 liposarcoma
#nonLPS : 3. Leiomyosarcoma / 4. MPNST / 5. Solitary fibrous tumor / 6. PEComa / 7. 중 comment 에 liposarcoma 없음.

out$TumorHistologicSubtype <- as.integer((c[["병리결과\r\n\r\n0. WD Liposarcoma\r\n1. DD Liposarcoma\r\n2. Pleomorphic Liposarcoma\r\n3. Leiomyosarcoma\r\n4. MPNST\r\n5. Solitary fibrous tumor\r\n6. PEComa\r\n7. Other"]] %in% c(0, 1, 2)) |
                                       ((c[["병리결과\r\n\r\n0. WD Liposarcoma\r\n1. DD Liposarcoma\r\n2. Pleomorphic Liposarcoma\r\n3. Leiomyosarcoma\r\n4. MPNST\r\n5. Solitary fibrous tumor\r\n6. PEComa\r\n7. Other"]] == 7) &
                             grepl("liposarcoma|Liposarcoma", c[["Other \r\n\r\ncomment"]])))  

#FNCLCC grade
out$FNCLCC <- as.factor(c[["FNCLCC grade\r\n\r\n1. total score 2-3\r\n2. total score 4-5\r\n3. total score 6,7,8"]])

#Tumor Resection
#R0/R1="0", R2="1", other=NA ("2"도 NA에 포함)
out$Resection <- c[["Surgical margins\r\n\r\n0. R0/R1\r\n1. R2\r\n2. Not available"]]
out$Resection <- as.factor(ifelse(out$Resection=="2", NA, out$Resection))

#Combined Organ Resection
# "colon resection" : Rt. + Lt. + rectum 
# "small bowel resection" : small bowel + duodenum
# "pancreas resection" : distal pan + PD 
# "liver resection" 
# "major vessel resection" : iliac a & v, IVC, aorta
out$Resection_Colon <- as.integer(
  (c[["동반절제 장기\r\nRight colon\r\n\r\n0. No\r\n1. Yes"]] == "1") | (c[["동반절제 장기\r\nLeft colon\r\n\r\n0. No\r\n1. Yes"]] == "1") | (c[["동반절제 장기\r\nRectum\r\n\r\n0. No\r\n1. Yes"]] == "1")
  )

out$Resection_SmallBowel <- as.integer(
  (c[["동반절제 장기\r\nSmall bowel\r\n\r\n0. No\r\n1. Yes"]] == "1") | (c[["동반절제 장기\r\nDuodenum\r\n\r\n0. No\r\n1. Yes"]] == "1")
  )

out$Resection_Pancreas <- as.integer(
  (c[["동반절제 \r\n장기\r\nDistal pancreas\r\n\r\n0. No\r\n1. Yes"]] == "1") | (c[["동반절제 \r\n장기\r\nPanreatico-duodenectomy\r\n\r\n0. No\r\n1. Yes"]] == "1")
  )

out$Resection_Liver <- as.integer((c[["동반절제 장기\r\nLiver\r\n\r\n0. No\r\n1. Yes"]]=="1"))

out$Resection_MajorVesselResection <- as.integer((c[["동반절제 장기\r\nIliac vein\r\n\r\n0. No\r\n1. Yes"]] == "1") |
                                                   (c[["동반절제 장기\r\nIVC\r\n\r\n0. No\r\n1. Yes"]] == "1") |
                                                   (c[["동반절제 장기\r\nIliac artery\r\n\r\n0. No\r\n1. Yes"]] == "1") |
                                                   (c[["동반절제 장기\r\nAorta\r\n\r\n0. No\r\n1. Yes"]] == "1"))

#OP time
out$opTime <- as.numeric(c[["수술시간\r\n(min)"]])

#intra OP transfusion
out$intraOpTransfusion <- as.integer(c[["PRBC 수혈 수"]])

#Estimated blood loss
out$EBL <- as.numeric(c[["EBL\r\n(ml)"]])

#Follow up duration
out$day_FU <- as.numeric(c[["마지막 f/u\r\n\r\ndd-mm-yyyy"]] - c[["수술날짜\r\n\r\ndd-mm-yyyy"]])

out
```

Data for Table2
```{r}
#Clavien Dindo grade
#complication이 TRUE여야 grade가 2~5 생기는데 grade가 0인 경우는 사실 complication false에 해당

out$ClavienDindoComplicationYesNo <- as.integer(c[["Clavien-Dindo complication \r\n\r\n0. No\r\n1. Yes"]])
out$ClavienDindoGrade <- c[["Clavien-Dindo grade \r\n\r\n2/3a/3b/4a/4b/5"]]
out$ClavienDindoGrade <- as.factor(ifelse(out$ClavienDindoGrade== "0" , NA, out$ClavienDindoGrade))
out$ClavienDindoComplicationSevere<-ifelse(out$ClavienDindoComplicationYesNo==1 & c[["Clavien-Dindo grade \r\n\r\n2/3a/3b/4a/4b/5"]]=="2",
                                     0,out$ClavienDindoComplicationYesNo)

#post OP transfusion
out$postOpTransfusion <- as.integer(c[["수술 후 PRBC 수혈 여부\r\n\r\n0. No\r\n1. Yes"]])

#ICU care
out$ICUcare <- as.integer(c[["ICU 입실여부\r\n\r\n0. No\r\n1. Yes"]])

#Return to OR
out$ReOP <- as.integer(c[["합병증으로 인한 Re-op 여부\r\n\r\n0. No\r\n1. Yes"]])

#Hospital Stay after OP
out$HospitalDay <- as.numeric(c[["재원일수(days)"]])

# Complications
for (vname in names(c)[c(128:135, 137:139)]){
  vn.new <- gsub(" ", ".", strsplit(vname, "\r")[[1]][1])
  out[[vn.new]] <- as.integer(c[[vname]])
}
out$Complication.others<-as.integer(ifelse(is.na(c[["\r\nOthers\r\n\r\ncomment"]]),0,1))
out
```

Data for Table3
```{r}
#RTx gray
out$RTgray <- c[["RT dose\r\n(Gy)"]]

out

```

Data for Table4,5
```{r}
#Long term outcomes
#Death : Alive(0) : FALSE, 사망(1) : TRUE, UK(2) : NA에 포함
out$Death <- ifelse(c[["사망여부\r\n\r\n0.Alive\r\n1.Dead\r\n2.Unknown.y"]] == "1", 1, ifelse(c[["사망여부\r\n\r\n0.Alive\r\n1.Dead\r\n2.Unknown.y"]]== "0", 0, NA))

#Recurrence
out$recur_local <- c[["재발#1\r\n\r\n0: 무\r\n1: 유.x"]]
out$recur_site <- c$`Site of local recurrence`
out$recur_site <- ifelse(out$recur_site == "6", NA, out$recur_site)
out$recur_day <- ifelse(c[["재발#1\r\n\r\n0: 무\r\n1: 유.x"]] == 1, 
                        as.numeric(as.Date(as.integer(c[["Date of local recurrence"]]), origin = "1899-12-30") - as.Date(c[["수술날짜\r\n\r\ndd-mm-yyyy"]])),
                        as.numeric(c[["마지막 f/u\r\n\r\ndd-mm-yyyy"]] - c[["수술날짜\r\n\r\ndd-mm-yyyy"]]))

out
```

### Table1 : Patients and operation characteristics
```{r}

my_vars<-names(out)
factor_vars <- c(names(out)[sapply(out, function(x){length(table(x))}) <= 5])
conti_vars <- setdiff(my_vars, factor_vars)
for (v in factor_vars){out[[v]] <- factor(out[[v]])}

library(tableone)

table1_vars<-c("Age","Sex","BMI","BMI_cat","DM","HTN","COPD","CoronaryArteryDisease","ChronicRenalDisease",
               "PrevAbdominalOp","preOpChemo","Hb","Hb_below9","Hb_below10","Albumin","Albumin_below3","PLT","PLT_below50","PLT_below100",
               "PT_INR","PT_INR_over1.5","PrimaryRecurrent","TumorSize","TumorHistologicSubtype","FNCLCC",
               "Resection","Resection_Colon","Resection_SmallBowel","Resection_Pancreas","Resection_Liver","Resection_MajorVesselResection",
               "opTime","intraOpTransfusion","EBL","day_FU")

table1_factor_vars<-table1_vars[table1_vars %in% factor_vars]
table1_conti_vars<-table1_vars[table1_vars %in% conti_vars]

#Group 1 vs 2 vs 3
vars.fisher <- sapply(table1_factor_vars,
                      function(x){is(tryCatch(chisq.test(table(out[["Group"]], out[[x]])),error = function(e) e, warning=function(w) w), "warning")})
vars.fisher <- table1_factor_vars[vars.fisher]
table1_Group <- CreateTableOne(vars=table1_vars, strata = "Group", data=out, factorVars = table1_factor_vars, argsNormal = list(var.equal = F))
print(table1_Group, showAllLevels = TRUE, exact = vars.fisher)

#Group 1 vs 23
vars.fisher <- sapply(table1_factor_vars,
                      function(x){is(tryCatch(chisq.test(table(out[["Group1vs23"]], out[[x]])),error = function(e) e, warning=function(w) w), "warning")})
vars.fisher <- table1_factor_vars[vars.fisher]
table1_Group1vs23 <- CreateTableOne(vars=table1_vars, strata = "Group1vs23", data=out, factorVars = table1_factor_vars, argsNormal = list(var.equal = F))
print(table1_Group1vs23, showAllLevels = TRUE, exact = vars.fisher)

#Group 1 vs 2
vars.fisher <- sapply(table1_factor_vars,
                      function(x){is(tryCatch(chisq.test(table(out[["Group1vs2"]], out[[x]])),error = function(e) e, warning=function(w) w), "warning")})
vars.fisher <- table1_factor_vars[vars.fisher]
table1_Group1vs2 <- CreateTableOne(vars=table1_vars, strata = "Group1vs2", data=out, factorVars = table1_factor_vars, argsNormal = list(var.equal = F))
print(table1_Group1vs2, showAllLevels = TRUE, exact = vars.fisher)

```

### Table2 : Complication outcomes including mortality
```{r}

table2_vars<-c("ClavienDindoComplicationYesNo","ClavienDindoGrade","ClavienDindoComplicationSevere",
               "postOpTransfusion","ICUcare","ReOP","HospitalDay")
complication_vars<-c("Abdominal.abscess","Bowel.anastomosis.leak","Biliary.leak","Bleeding","Evisceration","DVT",
                     "Lymphatic.leak","Pancreatic.leak","Sepsis","Urinary.leak","Ileus","Complication.others")

table2_factor_vars<-table2_vars[table2_vars %in% factor_vars]
table2_conti_vars<-table2_vars[table2_vars %in% conti_vars]


#Group 1 vs 2 vs 3
vars.fisher <- sapply(table2_factor_vars,
                      function(x){is(tryCatch(chisq.test(table(out[["Group"]], out[[x]])),error = function(e) e, warning=function(w) w), "warning")})
vars.fisher <- table2_factor_vars[vars.fisher]
table2_Group <- CreateTableOne(vars=table2_vars, strata = "Group", data=out, factorVars = table2_factor_vars, argsNormal = list(var.equal = F))
print(table2_Group, showAllLevels = TRUE, exact = vars.fisher)

#Group 1 vs 23
vars.fisher <- sapply(table2_factor_vars,
                      function(x){is(tryCatch(chisq.test(table(out[["Group1vs23"]], out[[x]])),error = function(e) e, warning=function(w) w), "warning")})
vars.fisher <- table2_factor_vars[vars.fisher]
table2_Group1vs23 <- CreateTableOne(vars=table2_vars, strata = "Group1vs23", data=out, factorVars = table2_factor_vars, argsNormal = list(var.equal = F))
print(table2_Group1vs23, showAllLevels = TRUE, exact = vars.fisher)

#Group 1 vs 2
vars.fisher <- sapply(table2_factor_vars,
                      function(x){is(tryCatch(chisq.test(table(out[["Group1vs2"]], out[[x]])),error = function(e) e, warning=function(w) w), "warning")})
vars.fisher <- table2_factor_vars[vars.fisher]
table2_Group1vs2 <- CreateTableOne(vars=table2_vars, strata = "Group1vs2", data=out, factorVars = table2_factor_vars, argsNormal = list(var.equal = F))
print(table2_Group1vs2, showAllLevels = TRUE, exact = vars.fisher)


```

### Table3 : Radiation gray among groups
```{r}
#Group 1 vs 2 vs 3
table3_Group <- CreateTableOne(var="RTgray",strata="Group",data=out,argsNormal = list(var.equal = F))
table3_Group

#Group 1 vs 23
table3_Group1vs23 <- CreateTableOne(var="RTgray",strata="Group1vs23",data=out,argsNormal = list(var.equal = F))
table3_Group1vs23

#Group 1 vs 2
table3_Group1vs2 <- CreateTableOne(var="RTgray",strata="Group1vs2",data=out,argsNormal = list(var.equal = F))
table3_Group1vs2

#Total table
table3_total<-c(as.integer(summary(out$Group)[1]),as.integer(summary(out$Group)[2]),as.integer(summary(out$Group)[3]),"","","")

p <- summary(aov(RTgray~Group, data = out, var.equal = F))[[1]][["Pr(>F)"]][1]
p1vs23 <- t.test(RTgray~Group1vs23, data = out, var.equal = F)$p.value
p1vs2 <- t.test(RTgray~Group1vs2, data = out, var.equal = F)$p.value
mean_sd <- aggregate(RTgray~Group, data = out, FUN = function(x){c(mean = mean(x), sd = sd(x))})

table3_total<-rbind(table3_total,
                    c(as.character(c(paste0(round(mean_sd[["RTgray"]][, "mean"], 2), " (", round(mean_sd[["RTgray"]][, "sd"], 2), ")"))),
                      ifelse(p < 0.001, "< 0.001", round(p, 3)),
                      ifelse(p1vs23 < 0.001, "< 0.001", round(p1vs23, 3)),
                      ifelse(p1vs2 < 0.001, "< 0.001", round(p1vs2, 3))))

colnames(table3_total)<-c("Group1","Group2","Group3","p-value(all)","p-value(1vs23)","p-value(1vs2)")
rownames(table3_total) <- c("n","RTgray")

table3_total

```

### Table4 : M+M risk factor analysis
Recur KM plots
```{r}
library(jskm)
library(survival)
recurfit_Group<-survfit(Surv(recur_day, as.integer(as.vector(recur_local))) ~ Group, data = out)
jskm(recurfit_Group, timeby=60, cumhaz = T, mark=F, table = T, xlims=c(0,720), ylims=c(0,0.6))

recurfit_Group1vs23<-survfit(Surv(recur_day, as.integer(as.vector(recur_local))) ~ Group1vs23, data = out)
jskm(recurfit_Group1vs23, timeby=60, cumhaz = T, mark=F, table = T, xlims=c(0,720), ylims=c(0,0.6))

recurfit_Group1vs2<-survfit(Surv(recur_day, as.integer(as.vector(recur_local))) ~ Group1vs2, data = out)
jskm(recurfit_Group1vs2, timeby=60, cumhaz = T, mark=F, table = T, xlims=c(0,720), ylims=c(0,0.6))
```

Death KM plots
```{r}
Deathfit_Group<-survfit(Surv(day_FU, as.integer(as.vector(Death))) ~ Group, data = out)
jskm(Deathfit_Group, timeby=60, cumhaz = T, mark=F, table = T, xlims=c(0,720), ylims=c(0,0.6))

Deathfit_Group1vs23<-survfit(Surv(day_FU, as.integer(as.vector(Death))) ~ Group1vs23, data = out)
jskm(Deathfit_Group1vs23, timeby=60, cumhaz = T, mark=F, table = T, xlims=c(0,720), ylims=c(0,0.6))

Deathfit_Group1vs2<-survfit(Surv(day_FU, as.integer(as.vector(Death))) ~ Group1vs2, data = out)
jskm(Deathfit_Group1vs2, timeby=60, cumhaz = T, mark=F, table = T, xlims=c(0,720), ylims=c(0,0.6))
```

### Table5 : R2 resection risk factor analysis
```{r}
```
