---
title: "SarcomaPreOpRTx"
author: "Heejoo"
date: "`r Sys.Date()`"
output:
  html_document: 
    dev: jpeg
    df_print: paged
    fig_height: 6
    fig_width: 10
    highlight: textmate
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: no
    code_folding: show
  word_document:
    toc: yes
    toc_depth: '3'
  bookdown::word_document2:
    fig_caption: yes
editor_options:
  chunk_output_type: console
always_allow_html: yes
---

<style>
   p {line-height: 2.5em;}
   pre, code {white-space:pre !important; overflow-x:scroll !important}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

## Load & Input
```{r, echo=FALSE}
library(readxl)
library(dplyr)
library(tableone)
library(DT)

#read data

setwd("C:/Users/USER/Desktop/2020-1학기/의학통계/sarcoma/preOP RTx")

## 읽을 때부터 UK를 결측치로 
a <- excel_sheets("sarcoma data sheet SMC 20200811.xlsx") %>% 
  lapply(function(x){read_excel("sarcoma data sheet SMC 20200811.xlsx",sheet=x, skip=2, na = c("UK"))})

b <- a[[1]] %>% 
  left_join(a[[2]], by = "환자번호") %>% left_join(a[[3]], by = "환자번호") %>% left_join(a[[4]], by = "환자번호") %>%
  left_join(a[[5]], by = "환자번호") %>% left_join(a[[6]], by = "환자번호") %>% left_join(a[[7]], by = "환자번호")

#Age 계산
b$Age <- as.numeric(b[["수술날짜\r\n\r\ndd-mm-yyyy.x"]] - b[["생년월일\r\n\r\ndd-mm-yyyy"]])/365.25

```

## Methods
```{r, echo=FALSE}
c <- b

out <- c %>% select(환자번호,Age,`성별\r\n\r\nM/F`)
names(out)[3] <- "Sex"; names(out)[1] <- "ID"
out$Sex <- as.factor(out$Sex)

out$primaryTumor <- ifelse(c[["Primary 수술여부\r\n\r\n0. Primary tumor\r\n1. Residual after incomplete resection\r\n2. Local recurrence.x"]] == 0, 1, 0)

#preOpRTx : 시행(1|5)=TRUE, 시행x(4)=FALSE, 그외=NA
out$preOpRTx <- ifelse(c[["RT timing\r\n\r\n0.None \r\n1.Preop only\r\n2. IORT only\r\n3.Preop + IORT\r\n4.Postop only\r\n5.Preop + postop boost\r\n6.IORT + postop"]] %in% c("1", "5"),T,
                       ifelse(c[["RT timing\r\n\r\n0.None \r\n1.Preop only\r\n2. IORT only\r\n3.Preop + IORT\r\n4.Postop only\r\n5.Preop + postop boost\r\n6.IORT + postop"]] == "4",F, NA))

#TE : 삽입(1)=TRUE, 삽입x(0)=FALSE
out$TE <- ifelse(c[["Tisuue expander insertion \r\n유뮤\r\n\r\n0. No\r\n1. Yes"]] == "1",T,
                 ifelse(c[["Tisuue expander insertion \r\n유뮤\r\n\r\n0. No\r\n1. Yes"]]== "0",F, NA))

#Group
#"1" : preOP RTx 시행o
#"2" : preOP PTx 시행x, TE 삽입 x
#"3" : preOP RTx 시행x, TE 삽입 o
out$Group <- as.factor(ifelse(out$preOpRTx,"1",ifelse(out$TE==F,"2","3")))
out$Group1vs23 <- as.factor(ifelse(out$Group == 1,1,23))

out
```

Data for Table1
```{r, echo=FALSE}
#BMI
height <- as.numeric(c[["키\r\n(cm)"]])/100
weight <- as.numeric(c[["몸무게\r\n(kg)"]])
out$BMI <- weight/height/height
out$BMI_cat <- factor(ifelse(out$BMI < 18.5, "< 18.5", ifelse(out$BMI < 25, "< 25", ifelse(out$BMI < 30, "< 30", "≥ 30"))))

#DM : 1=TRUE, 0=FALSE
out$DM <- as.integer(c[["DM\r\n\r\n0. No\r\n1.yes"]])

#HTN : 1=TRUE, 0=FALSE
out$HTN <- as.integer(c[["HTN\r\n\r\n0. No\r\n1.yes"]])

#COPD : 1=TRUE, 0=FALSE
out$COPD <- as.integer(c[["COPD\r\n\r\n0. No\r\n1.yes"]])

#Coronary artery disease : 1=TRUE, 0=FALSE
out$CoronaryArteryDisease <- as.integer(c[["Coronary artery disease\r\n\r\n0. No\r\n1.yes"]])

#Chronic renal disease : 1=TRUE, 0=FALSE
out$ChronicRenalDisease <- as.integer(c[["Chronic renal disease\r\n\r\n0. No\r\n1.yes"]])

#Prev abdominal OP Hx : 0="무", 1="유", 2="기타(laparo)"
out$PrevAbdominalOp <- c[["이전\r\nabdominal op Hx \r\n여부\r\n\r\n0. 무\r\n1. 유\r\n2. 기타(laparo)"]]

#PreOP chemo : 1=TRUE, 0=FALSE
out$preOpChemo <- as.integer(c[["Neoadjuvant chemo 여부\r\n\r\n0.No\r\n1.Yes"]])

#Hb
out$Hb <- as.numeric(c[["수술전 \r\n피검사\r\n\r\nHb\r\n(g/dL)"]])
out$Hb_below9 <- as.integer(out$Hb < 9)  
out$Hb_below10 <- as.integer(out$Hb < 10)

#Albumin
out$Albumin <- as.numeric(c[["수술전 피검사\r\n\r\nAlbumin\r\n(g/dL)"]])
out$Albumin_below3 <- as.integer(out$Albumin < 3)

#PLT
out$PLT <- as.numeric(c[["수술전 피검사\r\n\r\nPlatelet\r\n(1000/uL)"]])
out$PLT_below50 <- as.integer(out$PLT < 50)
out$PLT_below100 <- as.integer(out$PLT < 100)

#PT INR
out$PT_INR <- as.numeric(c[["수술전 피검사\r\n\r\nPT(INR)"]])
out$PT_INR_over1.5 <- as.integer(out$PT_INR > 1.5)

#Tumor size
out$TumorSize <- as.numeric(c[["종양크기\r\nFirst dimension\r\n(mm)"]])

#Tumor histologic subtype
#LPS : 0. WD Liposarcoma / 1. DD Liposarcoma / 2. Pleomorphic Liposarcoma / 7. 중 comment 에 liposarcoma
#nonLPS : 3. Leiomyosarcoma / 4. MPNST / 5. Solitary fibrous tumor / 6. PEComa / 7. 중 comment 에 liposarcoma 없음.
out$Liposarcoma_postop <- as.integer((c[["병리결과\r\n\r\n0. WD Liposarcoma\r\n1. DD Liposarcoma\r\n2. Pleomorphic Liposarcoma\r\n3. Leiomyosarcoma\r\n4. MPNST\r\n5. Solitary fibrous tumor\r\n6. PEComa\r\n7. Other"]] %in% c(0, 1, 2)) |
                                       ((c[["병리결과\r\n\r\n0. WD Liposarcoma\r\n1. DD Liposarcoma\r\n2. Pleomorphic Liposarcoma\r\n3. Leiomyosarcoma\r\n4. MPNST\r\n5. Solitary fibrous tumor\r\n6. PEComa\r\n7. Other"]] == 7) &
                             grepl("liposarcoma|Liposarcoma", c[["Other \r\n\r\ncomment"]])))  

#FNCLCC tumor grade
out$FNCLCC_grade<-c[["FNCLCC grade\r\n\r\n1/2/3/UK"]]
out$FNCLCC_grade<-ifelse(out$FNCLCC_grade=="UK",NA,out$FNCLCC_grade)

out$FNCLCC_grade1 <- as.integer(out$FNCLCC_grade == 1)
out$FNCLCC_grade2 <- as.integer(out$FNCLCC_grade == 2)
out$FNCLCC_grade3 <- as.integer(out$FNCLCC_grade == 3)
out$FNCLCC_NA <-as.integer(is.na(out$FNCLCC_grade))

#Tumor Resection
#R0/R1="0", R2="1", other=NA ("2"도 NA에 포함)
out$Resection <- c[["Surgical margins\r\n\r\n0. R0/R1\r\n1. R2: post OP 1주 CT에서 있을시 포함,debulking op\r\n2. Not available"]]
out$Resection <- as.factor(ifelse(out$Resection=="2", NA, out$Resection))
out$Resection<-as.factor(ifelse(out$Resection=="0","R0/R1",ifelse(out$Resection=="1","R2",out$Resection)))

#Combined Organ Resection
out$resection_liver <- as.integer(c[["동반절제 장기\r\nLiver\r\n\r\n0. No\r\n1. Yes"]])
out$resection_largebowel <- as.integer(c[["동반절제 장기\r\nLeft colon\r\n\r\n0. No\r\n1. Yes"]] == 1
                        | c[["동반절제 장기\r\nRight colon\r\n\r\n0. No\r\n1. Yes"]] == 1
                        | c[["동반절제 장기\r\nRectum\r\n\r\n0. No\r\n1. Yes"]] == 1)
out$resection_uterus <- as.integer(c[["동반절제 장기\r\nUterus\r\n\r\n0. No\r\n1. Yes"]])
out$resection_kidney <- as.integer(c[["동반절제 장기\r\nKidney\r\n\r\n0. No\r\n1. Yes"]])
out$resection_spleen <- as.integer(c[["동반절제 장기\r\nspleen\r\n\r\n0. No\r\n1. Yes"]])
out$resection_pancreas <- as.integer(c[["동반절제 \r\n장기\r\nPanreatico-duodenectomy\r\n\r\n0. No\r\n1. Yes"]] == 1
                      | c[["동반절제 \r\n장기\r\nDistal pancreas\r\n\r\n0. No\r\n1. Yes"]] == 1)
out$resection_smallbowel <- as.integer(c[["동반절제 장기\r\nSmall bowel\r\n\r\n0. No\r\n1. Yes"]] == 1
                      | c[["동반절제 장기\r\nDuodenum\r\n\r\n0. No\r\n1. Yes"]] == 1)
out$resection_stomach <- as.integer(c[["동반절제 장기\r\nStomach\r\n\r\n0. No\r\n1. Yes"]])

out$num_resected_organ <- rowSums(select(out, grep("resection_", names(out), value = T)), na.rm = T)

out$resection_vascular <- as.integer(
  (c[["동반절제 장기\r\nIliac vein\r\n\r\n0. No\r\n1. Yes"]] == "1") | (c[["동반절제 장기\r\nIVC\r\n\r\n0. No\r\n1. Yes"]] == "1") | (c[["동반절제 장기\r\nIliac artery\r\n\r\n0. No\r\n1. Yes"]] == "1") | (c[["동반절제 장기\r\nAorta\r\n\r\n0. No\r\n1. Yes"]] == "1")
)

#OP time
out$opTime <- as.numeric(c[["수술시간\r\n(min)"]])

#intra OP transfusion
out$intraOpTransfusion <- as.integer(c[["PRBC 수혈 수"]])

#Estimated blood loss
out$EBL <- as.numeric(c[["EBL\r\n(ml)"]])

#PET　SUVmax 값 관련 변수 및 분석 추가 될 예정.

out[c(1,9:48)]
```

Data for Table2
```{r, echo=FALSE}
#Clavien Dindo grade
#complication이 TRUE여야 grade가 2~5 생기는데 grade가 0인 경우는 사실 complication false에 해당

out$ClavienDindoComplication01 <- as.integer(c[["Clavien-Dindo complication \r\n\r\n0. No\r\n1. Yes"]])
out$ClavienDindoComplication_wo_2<-ifelse(out$ClavienDindoComplication01==1 & c[["Clavien-Dindo grade \r\n\r\n2/3a/3b/4a/4b/5"]]=="2",
                                     0,out$ClavienDindoComplication01)
out$ClavienDindoGrade <- c[["Clavien-Dindo grade \r\n\r\n2/3a/3b/4a/4b/5"]]
out$ClavienDindoGrade <- as.factor(ifelse(out$ClavienDindoGrade== "0" | is.na(out$ClavienDindoGrade) , "1", out$ClavienDindoGrade))

#post OP transfusion
out$postOpTransfusion <- as.integer(c[["수술 후 PRBC 수혈 여부\r\n\r\n0. No\r\n1. Yes"]])

#ICU care
out$ICUcare <- as.integer(c[["ICU 입실여부\r\n\r\n0. No\r\n1. Yes"]])

#Return to OR
out$ReOP <- as.integer(c[["합병증으로 인한 Re-op 여부\r\n\r\n0. No\r\n1. Yes"]])

#Hospital Stay after OP
out$HospitalDay <- as.numeric(c[["재원일수(days)"]])

#PreOP RTx to OP duration
out$preOP_Rtx_to_OP_day <- c$`preOP RTx to OP duration`

# Complications
for (vname in names(c)[c(128:135, 137:139)]){
  vn.new <- gsub(" ", ".", strsplit(vname, "\r")[[1]][1])
  out[[vn.new]] <- as.integer(c[[vname]])
}
out$Complication.others<-as.integer(ifelse(is.na(c[["\r\nOthers\r\n\r\ncomment"]]),0,1))

out[c(1,49:68)]
```

Data for Table3
```{r, echo=FALSE}
#RTx gray
out$RTgray <- as.numeric(c[["RT dose\r\n(Gy)"]])

out[c("ID","RTgray")]
```

Data for Table4,5
```{r, echo=FALSE}
#Long term outcomes

#Follow up duration
out$day_FU <- as.numeric(c[["마지막 f/u\r\n\r\ndd-mm-yyyy"]] - c[["수술날짜\r\n\r\ndd-mm-yyyy.x"]])

#Death : Alive(0) : FALSE, 사망(1) : TRUE, UK(2) : NA에 포함
out$Death <- ifelse(c[["사망여부\r\n\r\n0.Alive\r\n1.Dead\r\n2.Unknown.y"]] == "1",1,
                    ifelse(c[["사망여부\r\n\r\n0.Alive\r\n1.Dead\r\n2.Unknown.y"]]== "0", 0, NA))

#Recurrence
out$recur_local <- c[["재발#1\r\n\r\n0: 무\r\n1: 유.x"]]
out$recur_site <- c$`Site of recurrence`
out$recur_site <- ifelse(out$recur_site == "6", NA, out$recur_site)
out$recur_day <- ifelse(c[["재발#1\r\n\r\n0: 무\r\n1: 유.x"]] == 1, 
                        as.numeric(as.Date(as.integer(c[["Date of local recurrence"]]), origin = "1899-12-30") - as.Date(c[["수술날짜\r\n\r\ndd-mm-yyyy.x"]])),
                        as.numeric(c[["마지막 f/u\r\n\r\ndd-mm-yyyy"]] - c[["수술날짜\r\n\r\ndd-mm-yyyy.x"]]))

out[c(1,70:74)]
```

Variable categorizing
```{r, echo=FALSE}
my_vars<-names(out)
factor_vars <- c(names(out)[sapply(out, function(x){length(table(x))}) <= 5],"ClavienDindoGrade")
conti_vars <- setdiff(my_vars, factor_vars)
for (v in factor_vars){out[[v]] <- factor(out[[v]])}

#factor_vars 중 fisher test 해야하는 변수들
vars.fisher <- sapply(factor_vars,
                      function(x){is(tryCatch(chisq.test(table(out[["Group"]], out[[x]])),error = function(e) e, warning=function(w) w), "warning")})
vars.fisher <- factor_vars[vars.fisher]
```

### Table1 : Patients and operation characteristics
```{r, echo=FALSE}
table1_vars<-c("primaryTumor","Age","Sex","BMI","BMI_cat","DM","HTN","COPD","CoronaryArteryDisease","ChronicRenalDisease",
               "PrevAbdominalOp","preOpChemo","Hb","Hb_below9","Hb_below10","Albumin","Albumin_below3",
               "PLT","PLT_below50","PLT_below100","PT_INR","PT_INR_over1.5",
               "TumorSize","Liposarcoma_postop","FNCLCC_grade",
               "Resection","resection_liver","resection_largebowel","resection_uterus","resection_kidney",
               "resection_spleen","resection_pancreas","resection_smallbowel",
               "resection_stomach","resection_vascular",
               "opTime","intraOpTransfusion","EBL","day_FU")

table1_factor_vars<-table1_vars[table1_vars %in% factor_vars]
table1_conti_vars<-table1_vars[table1_vars %in% conti_vars]


table1<- lapply(table1_vars, function(va){
  if(va=="FNCLCC_grade"){
    tb <- table(out[[va]], out[["Group"]])
    p <- ifelse(va %in% vars.fisher, fisher.test(tb)$p.value, chisq.test(tb)$p.value)
    
    out.final <- cbind(paste0(va, " (%)"),"","","","",ifelse(p < 0.001, "< 0.001", round(p, 3)))
    
    tb<-table(out[["FNCLCC_grade1"]],out[["Group"]])
    tb.prop <- round(100 * prop.table(tb, 2), 2)      ## prop.table : 1- byrow 2 - bycol
    tb.out <- matrix(paste0(tb, " (", tb.prop, ")"), ncol =3)
    p <- ifelse(va %in% vars.fisher, fisher.test(tb)$p.value, chisq.test(tb)$p.value)
    out.final <-rbind(out.final,
                      c("","Grade 1",tb.out[2,],ifelse(p < 0.001, "< 0.001", round(p, 3))))
    
    tb<-table(out[["FNCLCC_grade2"]],out[["Group"]])
    tb.prop <- round(100 * prop.table(tb, 2), 2)      ## prop.table : 1- byrow 2 - bycol
    tb.out <- matrix(paste0(tb, " (", tb.prop, ")"), ncol =3)
    p <- ifelse(va %in% vars.fisher, fisher.test(tb)$p.value, chisq.test(tb)$p.value)
    out.final <-rbind(out.final,
                      c("","Grade 2",tb.out[2,],ifelse(p < 0.001, "< 0.001", round(p, 3))))
    
    tb<-table(out[["FNCLCC_grade3"]],out[["Group"]])
    tb.prop <- round(100 * prop.table(tb, 2), 2)      ## prop.table : 1- byrow 2 - bycol
    tb.out <- matrix(paste0(tb, " (", tb.prop, ")"), ncol =3)
    p <- ifelse(va %in% vars.fisher, fisher.test(tb)$p.value, chisq.test(tb)$p.value)
    out.final <-rbind(out.final,
                      c("","Grade 3",tb.out[2,],ifelse(p < 0.001, "< 0.001", round(p, 3))))
    
    tb<-table(out[["FNCLCC_NA"]],out[["Group"]])
    tb.prop <- round(100 * prop.table(tb, 2), 2)      ## prop.table : 1- byrow 2 - bycol
    tb.out <- matrix(paste0(tb, " (", tb.prop, ")"), ncol =3)
    out.final <-rbind(out.final,
                      c("","NA",tb.out[2,],""))
    
    return(out.final)
    
  } else if (va %in% table1_conti_vars){
    
    forms <- as.formula(paste0(va, "~ Group"))
    mean_sd <- aggregate(forms, data = out, FUN = function(x){c(mean = mean(x), sd = sd(x))})
    
    p<-oneway.test(forms, data = out)$p.value
    
    out.final <- c(va, "",
                   paste0(round(mean_sd[[va]][, "mean"], 2), " (", round(mean_sd[[va]][, "sd"], 2), ")"),
                   ifelse(p < 0.001, "< 0.001", round(p, 3)))
    
    return(out.final)
    
  } else if(va %in% table1_factor_vars) {
    
    tb <- table(out[[va]], out[["Group"]])
    tb.prop <- round(100 * prop.table(tb, 2), 2)      ## prop.table : 1- byrow 2 - bycol
    tb.out <- matrix(paste0(tb, " (", tb.prop, ")"), ncol =3)
    
    p <- ifelse(va %in% vars.fisher, fisher.test(tb)$p.value, chisq.test(tb)$p.value)
  
    if(nlevels(out[[va]])==2){
      if(levels(out[[va]])[[2]]=="1"){
        out.final <- c(paste0(va, " (%)"),"Yes",tb.out[2,],ifelse(p < 0.001, "< 0.001", round(p, 3)))
      }
      else{
        out.final <- cbind(c(paste0(va, " (%)"), rep("", nrow(tb.out) - 1)),
                           rownames(tb),
                           tb.out,
                           c(ifelse(p < 0.001, "< 0.001", round(p, 3)), rep("", nrow(tb.out) - 1)))
      }
    }else{
      out.final <- cbind(c(paste0(va, " (%)"), rep("", nrow(tb.out) - 1)),
                         rownames(tb),
                         tb.out,
                         c(ifelse(p < 0.001, "< 0.001", round(p, 3)), rep("", nrow(tb.out) - 1)))
    }
    return(out.final)
  }
}) %>% Reduce(rbind, .)

table1<-rbind(c("n","",as.integer(summary(out$Group))[1:3],""),table1)
colnames(table1) <- c("Variable", "Subgroup", "preOP RTx (+)", "preOP RTx (-) & TE (-)","preOP RTx (-) & TE (+)","p-value")
rownames(table1)<-NULL
datatable(table1)

```

### Table2 : Complication outcomes including mortality
Table2
```{r, echo=FALSE}
table2_vars<-c("ClavienDindoGrade","ClavienDindoComplication01","ClavienDindoComplication_wo_2",
               "postOpTransfusion","ICUcare","ReOP","HospitalDay")

table2_factor_vars<-table2_vars[table2_vars %in% factor_vars]
table2_conti_vars<-table2_vars[table2_vars %in% conti_vars]

table2<- lapply(table2_vars, function(va){
  if (va %in% table2_conti_vars){
    
    forms <- as.formula(paste0(va, "~ Group"))
    mean_sd <- aggregate(forms, data = out, FUN = function(x){c(mean = mean(x), sd = sd(x))})
    
    p<-oneway.test(forms, data = out)$p.value
    
    out.final <- c(va, "",
                   paste0(round(mean_sd[[va]][, "mean"], 2), " (", round(mean_sd[[va]][, "sd"], 2), ")"),
                   ifelse(p < 0.001, "< 0.001", round(p, 3)))
    
    return(out.final)
    
  } else if(va %in% table2_factor_vars) {
    
    tb <- table(out[[va]], out[["Group"]])
    tb.prop <- round(100 * prop.table(tb, 2), 2)      ## prop.table : 1- byrow 2 - bycol
    tb.out <- matrix(paste0(tb, " (", tb.prop, ")"), ncol =3)
    
    p <- ifelse(va %in% vars.fisher, fisher.test(tb)$p.value, chisq.test(tb)$p.value)
    
    if(nlevels(out[[va]])==2){
      if(levels(out[[va]])[[2]]=="1"){
        out.final <- c(paste0(va, " (%)"),"Yes",tb.out[2,],ifelse(p < 0.001, "< 0.001", round(p, 3)))
      }
      else{
        out.final <- cbind(c(paste0(va, " (%)"), rep("", nrow(tb.out) - 1)),
                           rownames(tb),
                           tb.out,
                           c(ifelse(p < 0.001, "< 0.001", round(p, 3)), rep("", nrow(tb.out) - 1)))
      }
    }else{
      out.final <- cbind(c(paste0(va, " (%)"), rep("", nrow(tb.out) - 1)),
                         rownames(tb),
                         tb.out,
                         c(ifelse(p < 0.001, "< 0.001", round(p, 3)), rep("", nrow(tb.out) - 1)))
    }
    return(out.final)
  }
}) %>% Reduce(rbind, .)

table2<-rbind(c("n","",as.integer(summary(out$Group))[1:3],"","",""),table2)
colnames(table2) <- c("Variable", "Subgroup", "preOP RTx (+)", "preOP RTx (-) & TE (-)","preOP RTx (-) & TE (+)","p-value")
rownames(table1)<-NULL
datatable(table2)
```
Figure 1
```{r, echo=FALSE}
library(ggplot2)
complication_vars<-c("Abdominal.abscess","Bowel.anastomosis.leak","Biliary.leak","Bleeding","Evisceration","DVT",
                     "Lymphatic.leak","Pancreatic.leak","Sepsis","Urinary.leak","Ileus","Complication.others")
for (v in complication_vars){out[[v]] <- as.integer(as.vector(out[[v]]))}

xx<-data.frame(out %>% select(`ClavienDindoComplication01`))
xx$ClavienDindoComplication01 <- reorder(xx$ClavienDindoComplication01,
                                          X = xx$ClavienDindoComplication01,
                                          FUN = function(x) -length(x))
xx$ClavienDindoComplication01<-factor(ifelse(xx$ClavienDindoComplication01=="0","No","Yes"))
label<-paste0(levels(xx$ClavienDindoComplication01),"\n",round(sort(table(xx))/sum(table(xx)),2) * 100,"%")

complication01.pie<-ggplot(xx, aes(x="", fill=ClavienDindoComplication01))+
  geom_bar(width = 1)+
  coord_polar("y",start=0.4)+
  theme(panel.grid=element_blank(),
        axis.text = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill="white"))+
  annotate(geom = "text", y = c(70,305), x = 1, label = label)

df<-data.frame(
  "Complication.Type"<-complication_vars,
  "Prop"<-matrix(colSums(out[complication_vars])/sum(rowSums(out[complication_vars])))
)
complication.type.pie<-ggplot(df, aes(x="", y=Prop, fill=Complication.Type))+ geom_bar(stat="identity", width=1)+
  geom_text(aes(label=paste0(round(Prop*100), "%"),x=1.2), color="white",
            position = position_stack(vjust = 0.5))+
  coord_polar("y", start=0)+
  ylab("Complication Proportions")+
  xlab("")+
  theme(panel.grid=element_blank(),axis.text = element_blank(),panel.background = element_rect(fill="white"))

complication01.pie
complication.type.pie

```

### Table3 : Radiation gray among groups
```{r, echo=FALSE}

table3_total<-c(as.integer(summary(out$Group))[1:3],"")

p<-oneway.test(RTgray~Group, data = out)$p.value
mean_sd <- aggregate(RTgray~Group, data = out, FUN = function(x){c(mean = mean(x), sd = sd(x))})

table3_total<-rbind(table3_total,
                    c(as.character(c(paste0(round(mean_sd[["RTgray"]][, "mean"], 2), " (", round(mean_sd[["RTgray"]][, "sd"], 2), ")"))),
                      ifelse(p < 0.001, "< 0.001", round(p, 3))))

colnames(table3_total)<-c("preOP RTx (+)", "preOP RTx (-) & TE (-)","preOP RTx (-) & TE (+)","p-value(all)")
rownames(table3_total) <- c("n","RTgray")

datatable(table3_total)

```

### Table4 : M+M risk factor analysis
Recur KM plots
```{r, echo=FALSE}
library(jskm)
library(survival)
recurfit_Group<-survfit(Surv(recur_day, as.integer(as.vector(recur_local))) ~ Group, data = out)
jskm(recurfit_Group, timeby=60, cumhaz = T, mark=F, table = T, xlims=c(0,720), ylims=c(0,0.6), main="Recurrence by group")
```

Death KM plots
```{r, echo=FALSE}
Deathfit_Group<-survfit(Surv(day_FU, as.integer(as.vector(Death))) ~ Group, data = out)
jskm(Deathfit_Group, timeby=60, cumhaz = T, mark=F, table = T, xlims=c(0,720), ylims=c(0,0.6), main="Death by group")
```

### Table5 : R2 resection risk factor analysis
```{r, echo=FALSE}
```
